---
title: "State of the art chapter - R code"
author: "Gard Jenset"
date: "Sunday, December 13, 2015"
output: html_document
---

This documents the statistical calculations on the state of the art analysis in chapter 2.

# The chi-square test of the full data

This test shows the relationship between corpus methods and quantitative methods in the historical-diachronic sample.

Build data matrix for chi-square:

```{r}
hd = matrix(c(33, 4, 11, 14), nrow=2)
rownames(hd) = c("Not corpus-based", "Corpus-based")
colnames(hd) = c("Qualitative", "Quantitative")
hd
```

Perform chi-square test and get Cramer V:

```{r}
chisq.test(hd)
source("cramer.r")
cv.test(hd)
```

# Calculate 95% confidence intervals

This section calculates 95% confidence intervals for the two proportions of quantitative data from the two samples (0.4 in the historical-diachronic sample, and 0.8 for Language).

Define a function to calculate the confidence interval:

```{r}
# pbar: proportion, i.e. point estimate
# n: number of observations
conf_int = function(pbar, n){
  SE = sqrt(pbar*(1-pbar)/n) # standard error
  E = qnorm(.975)*SE  # error margin
  pbar + c(-E, E) # confidence interval
}
```

Calculate intervals:

```{r}
# Language sample:
conf_int(pbar=0.8, n=15)

# historical sample:
conf_int(pbar=0.4, n=62)

```

# Test difference

We can use prop.test to formally establish that the proportions are signficantly different:

```{r}
# x=25: the number of quantitative papers in the historical sample
# n=62: the size of the historical sample
# p=0.8: the estimate of quantitative methods from Language
prop.test(x=25, n=62, p=0.8)
```

# MCA analysis

This section reports the MCA analysis of journals and methods.

Read full data and process
```{r}
meta = read.table("metastudy assessments.csv", header=T,sep="\t")

# drop some rows that are not linguistic / historical
meta2 = meta[-c(6, 26 ),]
meta2 = droplevels(meta2)

# set colnames
colnames(meta2) = c("assessor","journal","volume","issue","pages","evidence","corpusbased","quantitative","frequencies_are_just_mentioned","annotation","enriched","size_of_dataset_is_defined","additional_variables","hypothesis_mentioned","type_or_token_frequencies","data_availability","wordlists","Stat.tech","absolute_frequencies","percentages","basic_null_hypothesis_tests","linear_regresssion","philogenetic_trees","PCA","Comment")

meta2$neutral = ifelse(meta2$Comment == "Grammar writing, not really linguistic" |
meta2$Comment == "Mainly methodological" | meta2$Comment == "methodological paper" | meta2$Comment == "Theoretical overivew" |
meta2$Comment == "Theoretical paper", T, F)

meta3 = meta2[meta2$neutral == F | is.na(meta2$neutral),]
meta3 = droplevels(meta3)
```

Perform MCA analysis:

```{r}
library(ca)
meta_journ.ca = mjca(meta3[,c(2, 7, 8) ])
summary(meta_journ.ca)
plot(meta_journ.ca, map="symbiplot", mass=T, xlab="Dim 1: 90.9%", ylab="Dim 2: 0.8%", pch=16, col=c("gray","gray"), main="MCA plot, methods and journals")
box()
```


# Proportion test of reduced historical sample

If exclude "Language Variation and Change" from the historical sample, we can repeat the prop.test from above and compare the proportions of **corpus data use**, and find that the two samples are still different:

```{r}
# x=11: the number of corpus-based papers in the historical sample (no LVC)
# n=50: the size of the historical sample (no LVC)
# p=0.4: the estimate of corpus-use from Language
prop.test(x=11, n=50, p=0.4)
```

# Confidence intervals around corpus-use estimates

Here we see that if we take the 0.4 estimate of corpus-use from the 15 article sample from Language as a point estimate, the confidence interval overlaps with that of the reduced historical sample. However, the confidence interval **distributions** are clearly different:

```{r}
# empty plotting canvas:
plot(x=c(0,0.8), y=c(0,4.5), type="n", xlab="Probability", ylab="Density")

# confidence interval language:
lang_ci = conf_int(pbar=0.4, n=15)
# confidence interval reduce historical sample:
hist_ci = conf_int(pbar=0.22, n=50)

# calculate probability densities:
density_language = density(seq(from=lang_ci[1], to=lang_ci[2], by=0.01))
density_hist = density(seq(from=hist_ci[1], to=hist_ci[2], by=0.01))

# add distributions to plot:
lines(density_language, lty=1, lwd=2, col="darkgray")
lines(density_hist, lty=2, lwd=2, col="darkgray")

# add legend:
legend("topright", inset=0.05, legend=c("Language", "Historical linguistics"), lty=c(1, 2), lwd=2, col="darkgray")

```

Write plot above to file:

```{r}
pdf("conf_int_corpus_based.pdf")
# empty plotting canvas:
plot(x=c(0,0.8), y=c(0,4.5), type="n", xlab="Probability", ylab="Density")

# add distributions to plot:
lines(density_language, lty=1, lwd=2, col="darkgray")
lines(density_hist, lty=2, lwd=2, col="darkgray")

# add legend:
legend("topright", inset=0.05, legend=c("Language", "Historical linguistics"), lty=c(1, 2), lwd=2, col="darkgray")
dev.off()

```

# Methods in the two samples

Create a barplot of different quantitative techniques used in the sample, broken down by whether or not the technique was used in Language Variation and Change.


```{r}
quant.papers = meta3[meta3$quantitative==T,]
quant.papers = droplevels(quant.papers)
quant.papers$lvc = ifelse(quant.papers$journal=="LangVarChange", T, F)
qp.freq = xtabs(~lvc + absolute_frequencies, quant.papers)
qp.prop = xtabs(~lvc + percentages, quant.papers)
qp.nhts = xtabs(~lvc + basic_null_hypothesis_tests, quant.papers)
qp.linmod = xtabs(~lvc + linear_regresssion, quant.papers)
qp.trees = xtabs(~lvc + philogenetic_trees, quant.papers)
qp.pca = xtabs(~lvc + PCA, quant.papers)

# bind together as matrix
methods.x = matrix(
              c(qp.linmod[1,2], 
              qp.linmod[2,2], 
              qp.freq[1,2], 
              qp.freq[2,2], 
              qp.prop[1,2], 
              qp.prop[2,2], 
              qp.nhts[1,2], 
              qp.nhts[2,2], 
              qp.trees[1,2], 
              qp.trees[2,2], 
              qp.pca[1,2],
              qp.pca[2,2]),
              nrow=2
              )
rownames(methods.x) = c("Other", "LVC")
colnames(methods.x) = c("Linear models", "Freq", "%", "NHT", "Trees", "PCA")
# barplot
barplot(methods.x, names.arg= colnames(methods.x), legend.text=rownames(methods.x))

pdf("methods_barplot_journals.pdf")
barplot(methods.x, names.arg= colnames(methods.x), legend.text=rownames(methods.x))
dev.off()
```


