---
title: "EC analysis"
author: "Gard Jenset"
date: "Sunday, December 11, 2016"
output: html_document
---

# Read data

```{r}
me.ex <- read.table("me_there_07122016.txt", header=T, sep="\t")

# create log-transformed version of tree depth
me.ex$logMaxDepth <- log10(me.ex$maxDepth)

# create re-scaled year variable
me.ex$msDate50 <- me.ex$msDate/50

dim(me.ex)
summary(me.ex)
```



# Show relationship between constructional strength and T1/Ø
```{r}
set.seed(123)
plot(jitter(me.ex$msDate, 25), jitter(me.ex$nextProb, 3), pch = as.numeric(me.ex$emptyExpletive)+16, col=as.numeric(me.ex$emptyExpletive)+40, ylab="Conditional probability of next element", xlab="MS date", main="")
legend("topright", inset=0.05, legend=c("There", "Ø"), pch=c(16,17), col=c("gray", "black"))

# write image to file
pdf("ME_nextprod_by_date.pdf")
set.seed(123)
plot(jitter(me.ex$msDate, 25), jitter(me.ex$nextProb, 3), pch = as.numeric(me.ex$emptyExpletive)+16, col=as.numeric(me.ex$emptyExpletive)+40, ylab="Conditional probability of next element", xlab="MS date", main="")
legend("topright", inset=0.05, legend=c("There", "Ø"), pch=c(16,17), col=c("gray", "black"))
dev.off()

```

# Show distribution of subjects over time

```{r}
cdplot(as.factor(emptyExpletive)~msDate, data = me.ex, yaxlabels=c("There", "Ø"), xlab="MS date", ylab="Existential subject", main="Existential subject realisation")

# write image to file
pdf("ME_cdplot_subject_by_date_2.pdf")
cdplot(as.factor(emptyExpletive)~msDate, data = me.ex, yaxlabels=c("There", "Ø"), xlab="MS date", ylab="Existential subject", main="Existential subject realisation")
dev.off()

```



# Show numeric variables by dialect

```{r}
library(lattice)
xyplot(v1prop ~ msDate | dialect, data = me.ex, main="V1") 
xyplot(v2prop ~ msDate | dialect, data = me.ex, main="V2")
xyplot(nextProb ~ msDate | dialect, data = me.ex, main="Context")
xyplot(emptyExpletive ~ jitter(msDate) | dialect, data = me.ex, pch=16)

```

# Show response by dialect

```{r}
dialect_s.x <- xtabs(~emptyExpletive + dialect, me.ex)
barplot(dialect_s.x[,c("East Midlands", "West Midlands", "Southern", "Kentish", "Northern")], beside=T, legend.text=c("There", "Ø"), main="Existential subject by dialect", names.arg=c("East\nMidlands", "West\nMidlands", "Southern", "Kentish", "Northern"))
# text(cex=1, x=bx, y=-125, c("East Midlands", "West Midlands", "Southern", "Kentish", "Northern"), xpd=TRUE, srt=45)

pdf("barplot_dialect_subject.pdf")
barplot(dialect_s.x[,c("East Midlands", "West Midlands", "Southern", "Kentish", "Northern")], beside=T, legend.text=c("There", "Ø"), main="Existential subject by dialect", names.arg=c("East\nMidlands", "West\nMidlands", "Southern", "Kentish", "Northern"))
dev.off()
```

# Show response by genre

```{r}
genre_s.x <- xtabs(~emptyExpletive + genre, me.ex)
barplot(genre_s.x[,c("Religious Treatise", "History", "Homily", "Romance", "Travelogue", "Sermon", "Rule", "Fiction", "Biography, Life of Saint", "Bible", "Handbook", "Philosophy")], beside=T, legend.text=c("There", "Ø"), main="Existential subject by genre", names.arg=c("Religious\nTreatise", "History", "Homily", "Romance", "Travelogue", "Sermon", "Rule", "Fiction", "Biography", "Bible", "Handbook", "Philosophy"), las=2)


pdf("barplot_genre_subject.pdf")
barplot(genre_s.x[,c("Religious Treatise", "History", "Homily", "Romance", "Travelogue", "Sermon", "Rule", "Fiction", "Biography, Life of Saint", "Bible", "Handbook", "Philosophy")], beside=T, legend.text=c("There", "Ø"), main="Existential subject by genre", names.arg=c("Religious\nTreatise", "History", "Homily", "Romance", "Travelogue", "Sermon", "Rule", "Fiction", "Biography", "Bible", "Handbook", "Philosophy"), las=2)
dev.off()
```

# Try a simple logistic regression model as a baseline

Load packages and scripts:

```{r}
library(rms)
library(arm)
# script for Nagelkerke's R2
source("C:\\Users\\gbj\\Documents\\BACKUP\\R\\pseudoR2_2-0.r")
```


Scale proportion data:

```{r}
me.ex$nextProb01 <- me.ex$nextProb/0.1
me.ex$v1prop01 <- me.ex$v1prop/0.1
me.ex$v2prop01 <- me.ex$v2prop/0.1

```

Data without NA's:

```{r}
me.ex.nona <- na.omit(me.ex)
me.ex.nona <- droplevels(me.ex.nona)
dim(me.ex.nona)
nrow(me.ex) - nrow(me.ex.nona)  # this should equal 67
```


```{r}
fit0 <- lrm(emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 + v2prop01 + dialect + genre + msDate50, data = me.ex.nona)
fit0
fit0b <- glm(emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 + v2prop01 + dialect + genre + msDate50, data = me.ex.nona, family="binomial")
binnedplot(fitted(fit0b), resid(fit0b), main="fit 0 (glm)")
summary(fit0b)
```


# Do an lmer analysis



The model below converges, but the binned residuals plot indicates that the fit isn't great. However, the C-index is 0.85. The full model against a mixed intercept model gives a Nagelkerke R^2 of 0.99.


```{r}
# not converging:
# glmer(emptyExpletive ~  nextProb + v1prop + v2prop + mainV1 + mainV2 + period + (1 | dialect) + (1 | genre ), data = me.ex, family=binomial)
# ---
# glmer(emptyExpletive ~  nextProb + v1prop + v2prop + mainV1 + mainV2 + msDate + (1 | dialect) + (1 | genre ), data = me.ex, family=binomial)
# ---
# glmer(emptyExpletive ~  nextProb + logMaxDepth + v1prop + v2prop + mainV1 + mainV2 + genre + (1 | dialect) + (1 | period ), data = me.ex.nona, family=binomial)

fit1 <- glmer(emptyExpletive ~  nextProb + logMaxDepth + v1prop + v2prop + mainV1 + mainV2 + (1 | dialect) + (1 | period ) + (1|genre), data = me.ex.nona, family=binomial)

summary(fit1)

# check fit:
binnedplot(fitted(fit1), resid(fit1), main="Fit 1")

# Pseudo-R^2
# lmer.r2(f = c("emptyExpletive ~  nextProb + logMaxDepth + v1prop + v2prop + mainV1 + mainV2", "(1 | dialect) + (1 | period ) + (1|genre)"), d=me.ex.nona)

# C-index
p1 = 1/(1+exp(-fitted(fit1)))
somers2(p1, as.numeric(me.ex.nona$emptyExpletive))

```


Try simplifying the data variable:

```{r}
me.ex.nona$before1230 <- ifelse(me.ex.nona$msDate <= 1230, TRUE, FALSE)
summary(me.ex.nona$over1230)
tapply(me.ex.nona$msDate, me.ex.nona$before1230, mean)
```

Fit a slightly simplified model:

```{r}
fit2 <- glmer(emptyExpletive ~  nextProb + logMaxDepth + v1prop + v2prop + mainV1 + mainV2 + before1230 + (1 | dialect) + (1 | genre ), data = me.ex.nona, family=binomial)

summary(fit2)

# check fit:
binnedplot(fitted(fit2), resid(fit2), main="Fit 2")

# Pseudo-R^2
# lmer.r2(f = c("emptyExpletive ~  nextProb + v1prop + v2prop + mainV1 + mainV2 + before1230", "(1 | dialect) + (1 | genre )"), d=me.ex.nona)

# C-index
p2 = 1/(1+exp(-fitted(fit2)))
somers2(p2, as.numeric(me.ex.nona$emptyExpletive))
```

Try with interaction:

```{r}
fit3 <- glmer(emptyExpletive ~  nextProb + logMaxDepth + v1prop * mainV1 + v2prop + mainV2 + (1 | dialect) + (1 | period ) + (1|genre), data = me.ex.nona, family=binomial)

summary(fit3)

# check fit:
binnedplot(fitted(fit3), resid(fit3), main="Fit 3")

# Pseudo-R^2
# lmer.r2(f = c("emptyExpletive ~  nextProb + logMaxDepth + v1prop + v2prop + mainV1 + mainV2", "(1 | dialect) + (1 | period ) + (1|genre)"), d=me.ex.nona)

# C-index
p3 = 1/(1+exp(-fitted(fit3)))
somers2(p3, as.numeric(me.ex.nona$emptyExpletive))
```


```{r}
fit4a <- glmer(emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 * v2prop01 + (1 | dialect) + (1 | period ) + (1|genre), data = me.ex.nona, family=binomial)


fit4b <- glmer(emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 + v2prop01 + (1 | dialect) + (1 | period ) + (1|genre), data = me.ex.nona, family=binomial)

anova(fit4a, fit4b, test="Chi")
# there's a significant difference between the two -- 4a is better

summary(fit4a)

# check fit:
binnedplot(fitted(fit4a), resid(fit4a), main="Fit 4")

# Pseudo-R^2
lmer.r2(f = c("emptyExpletive ~  nextProb + logMaxDepth + I(v1prop/0.1) + I(v2prop/0.1)", "(1 | dialect) + (1 | period ) + (1|genre)"), d=me.ex.nona)

# C-index
p4 = 1/(1+exp(-fitted(fit4a)))
somers2(p4, as.numeric(me.ex.nona$emptyExpletive))
```

Try dialect and genre as a fixed effects:

```{r}
# 'genre' as fixed effect doesn't converge
# 'dialect' as fixed effect doesn't converge
fit5 <- glmer(emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 * v2prop01 + msDate + (1|genre) + (1 | dialect ), data = me.ex.nona, family=binomial)

anova(fit4a, fit5, test="Chi")
# there's no difference between the two = prefer the simpler fit5

summary(fit5)

# check fit:
binnedplot(fitted(fit5), resid(fit5), main="Fit 5")

# C-index
p5 = 1/(1+exp(-fitted(fit5)))
somers2(p5, as.numeric(me.ex.nona$emptyExpletive))

# Pseudo-R^2
lmer.r2(f = c("emptyExpletive ~  nextProb01 + logMaxDepth + v1prop01 * v2prop01", "(1 | period ) + (1|genre)"), d=me.ex.nona)


```



## calculate spread around intercept

```{r}
# for dialect
plogis(qnorm(c(0.025, 0.975), mean=summary(fit4a)$coefficients[1,1], sd=sqrt(summary(fit4a)$varcor$dialect)))

# for period
plogis(qnorm(c(0.025, 0.975), mean=summary(fit4a)$coefficients[1,1], sd=sqrt(summary(fit4a)$varcor$period)))

# for genre
plogis(qnorm(c(0.025, 0.975), mean=summary(fit4a)$coefficients[1,1], sd=sqrt(summary(fit4a)$varcor$genre)))
```



# Summary

None of the GLMM models can beat the basic glm model:

* C index is virtually identical
* residuals plot show better fit with the glm

```{r}
# generate table
source("binom_glmm_table_3.0.r")

# for inspection:
# table.glmm(fit4a, "md")
table.glmm(fit0b, "md")
```

For use in book:

```{r, comment=""}
# table.glmm(fit4a, "tex")
table.glmm(fit0b, "tex")
```

Write binned plot to file:

```{r}
pdf("binnedplot_glm0.pdf")
binnedplot(fitted(fit0b), resid(fit0b), main="Binned residuals")
dev.off()

```

