---
title: "s-th -- EME"
author: "Gard Jenset"
date: "Sunday, December 20, 2015"
output: html_document
---

```{r}
library(arm)
library(Hmisc)
# script for Nagelkerke's R2
# source("pseudoR2_2-0.r")

```

# Read data

```{r}
eme.raw = read.table('eme_3sg_data_15112015a.txt', header=T, sep='\t', quote="\"", fill=T)
eme.raw = eme.raw[eme.raw$filename != "later countess of Halifax",]
eme = na.omit(eme.raw)
eme = droplevels(eme)
eme$year = as.numeric(as.character(eme$year))
eme$yearGroup = cut2(eme$year, g=20, levels.mean=T)
eme$female = ifelse(eme$authorGender == "F", T, F)
eme$infs = ifelse(eme$suffix3sg == "s", T, F)
eme$lemma_len = nchar(as.character(eme$lemma))
```

# Plot over time

```{r}
s = eme[eme$suffix3sg == "s",]
s = droplevels(s)
th = eme[eme$suffix3sg == "th",]
th = droplevels(th)
par(mfrow=c(1,2))
plot(s$year, log10(s$subPeriodCount), col="grey", main="-s", xlab="Year", ylab="Verb frequency (base 10 log scale)")
lines(lowess(log10(s$subPeriodCount)~s$year), col="red", lwd = 3)
plot(th$year, log10(th$subPeriodCount), col="grey", main="-th", xlab="Year", ylab="Verb frequency (base 10 log scale)")
lines(lowess(log10(th$subPeriodCount)~th$year), col="red", lwd=3)

pdf("s-th_lowess_time.pdf")
par(mfrow=c(1,2))
plot(s$year, log10(s$subPeriodCount), col="grey", main="-s", xlab="Year", ylab="Verb frequency (base 10 log scale)")
lines(lowess(log10(s$subPeriodCount)~s$year), col="black", lwd = 3)
plot(th$year, log10(th$subPeriodCount), col="grey", main="-th", xlab="Year", ylab="Verb frequency (base 10 log scale)")
lines(lowess(log10(th$subPeriodCount)~th$year), col="black", lwd=3)
dev.off()

par(mfrow=c(1,1))
```
The plots show that the s form becomes more common over time (left plot), while the th form becomes less common over time (right plot), which we knew already.
In addition the left plot shows that for the s forms, there isn't much change associated with lemma frequency over time because both the frequent and the infrequent verbs adopt it increasingly over time.
However, the right plot shows that for the th forms there is change associated with lemma frequency over time because the th-form is becoming more and more limited to high-frequency verbs.

So we can conclude that we expect that over time the th form becomes more limited to high-frequency verbs.

# Plot S-curve

```{r}
cdplot(eme$suffix3sg ~ eme$year, xlab="Year", ylab="Suffix")

pdf("cdplot_suffix_time.pdf")
cdplot(eme$suffix3sg ~ eme$year, xlab="Year", ylab="Suffix")
dev.off()
```

# MCA

```{r}
library(ca)
library(dplyr)
eme.mca = mjca(eme[, c(11, 14, 2, 16)])
plot(eme.mca, map="rowgreen", xlab="Dimension 1: 71.5%", ylab="Dimension 2: 1.8%", col=c("black", "darkgray"))
summary(eme.mca)

pdf("s-th_MCA_plot.pdf")
plot(eme.mca, map="rowgreen", xlab="Dimension 1: 71.5%", ylab="Dimension 2: 1.8%", col=c("black", "darkgray"))
dev.off()

```

# Count female authors by period

```{r}
library(xtable)
fem.x = xtabs(~female+period, eme)
xtable(fem.x, digits=0)

prop.table(fem.x)

```

# 3 models by period

This is the best approach and the models we will properly report on.

## Get necessary functions

```{r}
library(arm)
library(rms)
# Script for generating summary table:
# source("binom_glmm_table_2.0.r")
# script for Nagelkerke's R2
# source("pseudoR2_2-0.r")

```

## Create the datasets

```{r}
period1 = eme[eme$period == "e1",]
period1 = droplevels(period1)

period2 = eme[eme$period == "e2",]
period2 = droplevels(period2)

period3 = eme[eme$period == "e3",]
period3 = droplevels(period3)
```

## Period E1

Create the model:
```{r}
glmer.p1 = glmer(infs ~ (1|genre) + I(log10(subPeriodCount))  + verbTag, data = period1, family = binomial)
 
binnedplot(fitted(glmer.p1), resid(glmer.p1), xlab="Estimated probability of \"-s\"")
summary(glmer.p1)

```

Generate the outputs below.

Summary table:
```{r}
table.glmm(glmer.p1)

```

Write binnedplot to file:
```{r}
pdf("E1_period_binned_plot.pdf")
binnedplot(fitted(glmer.p1), resid(glmer.p1), xlab="Estimated probability of \"-s\"")
dev.off()
```

Calculate R^2:
```{r}
lmer.r2(f = c("infs ~ I(log10(subPeriodCount)) + verbTag", "(1 | genre)"), d = period1)

```

Calculate Harrel's C:
```{r}
p.e1 = 1/(1+exp(-fitted(glmer.p1)))
somers2(p.e1, as.numeric(period1$infs))
```

## Period E2

Create the model:
```{r}

glmer.p2 = glmer(infs ~ (1|genre) + I(log10(subPeriodCount)) + female + context + verbTag, data = period2, family = binomial)
binnedplot(fitted(glmer.p2), resid(glmer.p2), xlab="Estimated probability of \"-s\"")
summary(glmer.p2)
```

Generate the outputs below.

Summary table:
```{r}
table.glmm(glmer.p2)

```

Write binnedplot to file:
```{r}
pdf("E2_period_binned_plot.pdf")
binnedplot(fitted(glmer.p2), resid(glmer.p2), xlab="Estimated probability of \"-s\"")
dev.off()
```

Calculate R^2:
```{r}
lmer.r2(f = c("infs ~ I(log10(subPeriodCount)) + female + context + verbTag", "(1 | genre)"), d = period2)

```

Calculate Harrel's C:
```{r}
p.e2 = 1/(1+exp(-fitted(glmer.p2)))
somers2(p.e2, as.numeric(period2$infs))
```

Calculate CI for random effect:

```{r}
# genre
plogis(qnorm(c(0.025, 0.975), mean=summary(glmer.p2)$coefficients[1,1], sd=sqrt(summary(glmer.p2)$varcor$genre)))
```



## Period E3

Create the model:
```{r}
glmer.p3 = glmer(infs ~ (1|genre) + I(log10(subPeriodCount)) + female + verbTag, data = period3, family = binomial)
binnedplot(fitted(glmer.p3), resid(glmer.p3), xlab="Estimated probability of \"-s\"")
summary(glmer.p3)
```

Generate the outputs below.

Summary table:
```{r}
table.glmm(glmer.p3)

```

Write binnedplot to file:
```{r}
pdf("E3_period_binned_plot.pdf")
binnedplot(fitted(glmer.p3), resid(glmer.p3), xlab="Estimated probability of \"-s\"")
dev.off()
```

Calculate R^2:
```{r}
lmer.r2(f = c("infs ~ I(log10(subPeriodCount)) + female + verbTag", "(1 | genre)"), d = period3)

```

Calculate Harrel's C:
```{r}
p.e3 = 1/(1+exp(-fitted(glmer.p3)))
somers2(p.e3, as.numeric(period3$infs))
```

Calculate CI for random effect:

```{r}
# genre
plogis(qnorm(c(0.025, 0.975), mean=summary(glmer.p3)$coefficients[1,1], sd=sqrt(summary(glmer.p3)$varcor$genre)))
```

# A bad model

```{r}

mod_bad= glmer(infs ~ (1|genre) + I(log10(subPeriodCount)) + period + female + context + verbTag, data = eme, family = binomial)

# binned resid plot
binnedplot(fitted(mod_bad), resid(mod_bad), xlab="Estimated probability of \"-s\"")

# Write binned plot file
pdf("bad_binned_plot.pdf")
binnedplot(fitted(mod_bad), resid(mod_bad), xlab="Estimated probability of \"-s\"")
dev.off()

summary(mod_bad)
```


Table output:
```{r}
table.glmm(mod_bad)

```

R^2:
```{r}
lmer.r2(f = c("infs ~ I(log10(subPeriodCount)) + period + female + context + verbTag", "(1 | genre)"), d = eme)
```

Harrel's C:
```{r}
p.bad = 1/(1+exp(-fitted(mod_bad)))
somers2(p.bad, as.numeric(eme$infs))
```
